---
title: "Simulated_case_study_6_6"
output: html_document
editor_options: 
  chunk_output_type: console
---

Simulation of two communities depicted in chapter 6 in JDSM

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulating species niches

### Construct random phylogeny

Phylogeny for 100 species converted into a phylogenetic correlation matrix.

```{r, echo = FALSE}

library(Hmsc)
library(ape)
library(MASS)
```

```{r, echo = FALSE}

ns <- 100
phy <- rcoal(n = ns, 
             tip.label = sprintf('sp_%.3d', 1:ns),
             br = "coalescent")
C <- vcv(phy, model = "Brownian", corr = TRUE)
# plot(phy)
```

## Sample trait values

-   Community A: Traits sampled independently from standard normal distribution

-   Community B: Traits sampled from multivariate normal distribution

    $\rightarrow$ variance-covariance matrix equals the phylogenetic correlation matrix

-   Intercept has been added

```{r, echo = FALSE}
Tr.A <- cbind(rep(1, ns), rnorm(ns))
Tr.B <- cbind(rep(1, ns), mvrnorm(n = 1, 
                                  mu = rep(0, ns),
                                  Sigma = C))
```

## Link species traits and niches

```{r, echo = FALSE}
gamma <- cbind(c(-2,2), c(-1,1))
mu.A <- gamma %*% t(Tr.A)
mu.B <- gamma %*% t(Tr.B)
```

## Generate species niches

Assumptions:

-   Community A: residual variation is **phylogenetically fully structured**

-   Community B: residual variation is **fully independent among species**

```{r}
V2 = diag(2)
beta.A <- matrix(mvrnorm(n = 1, 
                        mu = as.vector(mu.A), 
                        Sigma = kronecker(C, V2)), 
                ncol = ns)
beta.B <- matrix(mvrnorm(n = 1, 
                        mu = as.vector(mu.B), 
                        Sigma = kronecker(diag(ns), V2)), 
                ncol = ns)
```

## Simulate species data

Single environmental covariate:

```{r}
# include the intercept
n <- 50
X <- cbind(rep(1, n), rnorm(n))

# calculate linear predictors
L.A <- X %*% beta.A
L.B <- X %*% beta.B
```

Convert linear predictors into community data:

```{r}

# using the probit model for presences/absence 
Y.A <- 1*((L.A+matrix(rnorm(n*ns), ncol = ns)) > 0)
Y.B <- 1*((L.B+matrix(rnorm(n*ns), ncol = ns)) > 0)
```

## Explore the raw data

```{r}
# species richness
S.A <- rowSums(Y.A)
# species prevalences
P.A <- colMeans(Y.A)
S.B <- rowSums(Y.B)
P.B <- colMeans(Y.B)

# plot(X[,2], S.A)
# plot(X[,2], S.B)
# hist(P.A)
# hist(P.B)
```

## Date preparation & analysis

```{r}

# prep
community <-  "A"
Y <- switch(community, "A" = Y.A, "B" = Y.B)
colnames(Y) <- phy$tip.label
Tr <- switch(community, "A" = Tr.A, "B" = Tr.B)
TrData <- data.frame(trait = Tr[, 2])
XData <- data.frame(x = X[, 2])

# model
m <- Hmsc(Y = Y,
          XData = XData, 
          XFormula = ~ x, 
          TrData = TrData, 
          TrFormula = ~ trait, 
          phyloTree = phy, 
          distr = "probit")

# model fitting
nChains <- 2
thin <- 5
samples <- 1000
transient <- thin*500
verbose <- 500*thin
m = sampleMcmc(m, 
               thin = thin, 
               samples = samples,
               transient = 500*thin,
               nChains = nChains,
               verbose = verbose)

# extract posterior distribution 
mpost <- convertToCodaObject(m)

# evaluate MCMC convergence:
# effective size
effectiveSize(mpost$Rho)

# scale reduction factors
gelman.diag(mpost$Rho,
            multivariate = FALSE,
            autoburnin = FALSE)$psrf
```

## Explanatory and predictive power of HMSC model

```{r}
preds <- computePredictedValues(m)
MF <- evaluateModelFit(hM = m, 
                       predY = preds)

# with CV 
partition <- createPartition(m, nfolds = 2)
preds <- computePredictedValues(m, 
                                partition = partition)
MFCV <- evaluateModelFit(hM = m,
                         predY = preds)
```

## Examining parameter estimates 

Visualisation of estimated species niches

```{r}
postBeta <- getPostEstimate(m,
                            parName = "Beta")

# phylogenetic signal in the ß parameters
# entire clades show blue or red blocks 
plotBeta(m , 
         post = postBeta,
         param = "Sign",
         plotTree = TRUE, 
         supportLevel = 0.95, 
         split = 0.4,
         spNamesNumbers = c(F,F))

```

How species niches estimation depends on species traits?

```{r}
postGamma <- getPostEstimate(m, parName = "Gamma")
plotGamma(m, 
          post = postGamma,
          param = "Sign", 
          supportLevel = 0.95)
```

Parameter estimate for the phylogenetic signal

```{r}

# evidence for phylogenetic signal
# (95 % credible interval close to one)
summary(mpost$Rho)$quantiles
```

## Did including traits and phylogenies improve predictions?

Fit the same model to the same data, but without including information about traits and phylogeny.

```{r}
m <- Hmsc(Y = Y, 
          XData = XData,
          XFormula = ~x,
          distr = "probit")
m <- sampleMcmc(m, 
                thin = thin,
                samples = samples,
                transient = transient,
                nChains = nChains,
                verbose = verbose)

# model fit
preds <- computePredictedValues(m)
MF.NTP <- evaluateModelFit(hM = m, 
                           predY =  preds)

preds <- computePredictedValues(m, 
                                partition = partition)
MFCV.NTP <- evaluateModelFit(hM = m,
                             predY = preds)

Delta.TjurR2 <- MFCV$TjurR2 <- MFCV.NTP$TjurR2

# Species prevalence vs. Delta TjurR2
plot(P.A, Delta.TjurR2)
```

How well does estimated and true parameter values match?

```{r}
beta.slope.true <- beta.A[2, ]
beta.slope.est <- postBeta$mean[2, ]
postBeta.NTP <- getPostEstimate(m,
                                 parName = "Beta")
beta.slope.est.NTP <- postBeta.NTP$mean[2, ]

# comparison
plot(beta.slope.true, beta.slope.est)
plot(beta.slope.true, beta.slope.est.NTP)
```

## Re-run with community B... 
