---
title: "Simulated_case_study_6_6"
output: html_document
editor_options: 
  chunk_output_type: console
---

Simulation of two communities depicted in chapter 6 in JDSM

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulating species niches

### Construct random phylogeny

Phylogeny for 100 species converted into a phylogenetic correlation matrix.

```{r, echo = FALSE}
library(Hmsc)
library(ape)
library(MASS)
```

```{r, echo = FALSE}
ns <- 100
phy <- rcoal(n = ns, 
             tip.label = sprintf('sp_%.3d', 1:ns),
             br = "coalescent")
C <- vcv(phy, model = "Brownian", corr = TRUE)

plot(phy)
```

## Sample trait values

-   Community A: Traits sampled independently from standard normal distribution

-   Community B: Traits sampled from multivariate normal distribution

    $\rightarrow$ variance-covariance matrix equals the phylogenetic correlation matrix

-   Intercept has been added

```{r, echo = FALSE}
Tr.A <- cbind(rep(1, ns), rnorm(ns))
Tr.B <- cbind(rep(1, ns), mvrnorm(n = 1, 
                                  mu = rep(0, ns),
                                  Sigma = C))
```

## Link species traits and niches

```{r, echo = FALSE}
gamma <- cbind(c(-2,2), c(-1,1))
mu.A <- gamma %*% t(Tr.A)
mu.B <- gamma %*% t(Tr.B)
```

## Generate species niches

Assumptions:

-    Community A: residual variation is **phylogenetically fully structured**

-   Community B: residual variation is **fully independent among species**

```{r}
V2 = diag(2)
beta.A = matrix(mvrnorm(n = 1, 
                        mu = as.vector(mu.A), 
                        Sigma = kronecker(C, V2)), 
                ncol = ns)
beta.B = matrix(mvrnorm(n = 1, 
                        mu = as.vector(mu.B), 
                        Sigma = kronecker(diag(ns), V2)), 
                ncol = ns)
```

## Simulate species data 
